Fail Faster

### 2025-03-12 paper/paper.tex
blah

@@

\section{Why is \bq\ slow, and how can we fix it?}
\jwc{Why is a monadic generator library slow?}

@@

\section{\name: A Library for Staged Generators}

@@

This type, along with basic monadic generator DSL functionality can be found in Figure~\ref{fig:gen-staged-basic}.
The monadic interface is given by a return and bind, as usual.
\texttt{return} is the constant generator, but this time it operates on codes. Given \texttt{cx : 'a code}, the code for a \texttt{'a}, it returns the generator which always generates that value.
\texttt{bind g k} sequences generators by passing the result of running the generator \texttt{g} to a continuation \texttt{k}. However, instead of getting access
to the particular value generated by \texttt{g}, the continuation \texttt{k} gets access to \texttt{code} for the value sampled from \texttt{g}:
at compile time, we only know \texttt{g} will generate \emph{some} \texttt{'a}, but not which one.
Operationally, bind takes code for the size and seed, and returns code that (1) let-binds a variable \texttt{a} to spliced-in code that runs \texttt{g}, and then
(2) runs the spliced-in continuation \texttt{k}.
Both function applications \texttt{g size\_c random\_c}
and \texttt{k .<a>. size\_c random\_c} run at compile time.
\texttt{Gen.int} is the generator that samples an int from the RNG.
Given any size and random seed, it returns a code block that calls \texttt{SR.int} with that random seed. Because the lower and upper bounds
might not be known at compile time --- they may themselves be the results of calling \texttt{Gen.int} ---
the arguments \texttt{lo} and \texttt{hi} are of type \texttt{int code}, and get spliced into the code block as arguments to \texttt{SR.int}.
Lastly, \texttt{to_bq} turns a staged generator into code for a normal \bq
generator. This function is just a 2-argument version of ``The Trick'' (\texttt{eta}
from Section~\ref{subsection:msp}).

>>>>

\jwc{
  Things to ensure people come away with answers to:
  \begin{itemize}
    \item Is there generality?
    \item Is this important?
    \item Did they actually solve the problem?
  \end{itemize}
}

@@

\section{Why are Monadic Generator Libraries Slow?}
\jwc{Why is a monadic generator library slow?}

@@

\section{Eliminating Abstraction Overhead of Generator DSLs by Staging}
\jwc{Emphasize that this is an \emph{COMPLETELY EQUIVALENT drop in replacement!} We din't just build a different library with different distributions.}

@@

This type, along with basic monadic generator DSL functionality can be found in Figure~\ref{fig:gen-staged-basic}.
The monadic interface is given by a return and bind, as usual.
\texttt{return} is the constant generator, but this time it operates on codes. Given \texttt{cx : 'a code}, the code for a \texttt{'a}, it returns the generator which always generates that value.
\texttt{bind g k} sequences generators by passing the result of running the generator \texttt{g} to a continuation \texttt{k}. However, instead of getting access
to the particular value generated by \texttt{g}, the continuation \texttt{k} gets access to \texttt{code} for the value sampled from \texttt{g}:
at compile time, we only know \texttt{g} will generate \emph{some} \texttt{'a}, but not which one.
Operationally, bind takes code for the size and seed, and returns code that (1) let-binds a variable \texttt{a} to spliced-in code that runs \texttt{g}, and then
(2) runs the spliced-in continuation \texttt{k}.
Both function applications \texttt{g size\_c random\_c}
and \texttt{k .<a>. size\_c random\_c} run at compile time.
\texttt{Gen.int} is the generator that samples an int from the RNG.
Given any size and random seed, it returns a code block that calls \texttt{SR.int} with that random seed. Because the lower and upper bounds
might not be known at compile time --- they may themselves be the results of calling \texttt{Gen.int} ---
the arguments \texttt{lo} and \texttt{hi} are of type \texttt{int code}, and get spliced into the code block as arguments to \texttt{SR.int}.
Lastly, \texttt{to\_bq} turns a staged generator into code for a normal \bq
generator. This function is just a 2-argument version of ``The Trick'' (\texttt{eta}
from Section~\ref{subsection:msp}).
